---
title: "Analysing user interactions with the prototype"
output: html_notebook
---
In this notebook, I analyse how participants interacted with the prototype during the experiment. This includes analysing the usage of each of the prototype's functions.

## Include necessary packages and load data
```{r}
library(WRS2)
library(rcompanion)
library(feather)
library(ggplot2)  
library(ez)
library(schoRsch)
library(MASS)
library(car)
library(ggpubr)
library(rstatix)
library(tidyverse)
library(here)
```

```{r}
df <- read_feather(path = here('../../data_frames/results_test_02-04-2020_user_actions.file'))
```


## Prepare the data

### Assign participants to the experimental group 
* version 1 and 2 --> control group
* version 3 und 4 --> experimental group context free
* version 5 und 6--> experimental group context sensitive

```{r}
df$group <- ifelse((df$version == 5 | df$version == 6), 'context-sensitive experimental group', ifelse((df$version == 1 | df$version == 2), 'controlgroup', 'context-free experimental group'))
```

### Convert trial, group and subject number to factors in each dataframe
```{r}
df$trial <- factor(df$trial, levels = c("tutorial", "warm up", "trial 1", "trial 2", "trial 3"))
df$group <- factor(df$group)
df$partnr <- factor(df$partnr)
```


### subset df's 

```{r}
# df without tutorial and warm up tasks

df_trial <- df[!(df$trial=="tutorial" | df$trial=="warm up"), ]
```

```{r}
# user actions only for the control group

df_control <- df_trial[(df_trial$group=="controlgroup"), ]
```

```{r}
# user actions only for the context-free experimental group

df_exp1 <- df_trial[(df_trial$group=="context-free experimental group"), ]
```


```{r}
# user actions only for the context-sensitive experimental group

df_exp2 <- df_trial[(df_trial$group=="context-sensitive experimental group"), ]
```

```{r}
# user actions only for the experimental groups

df_exp <- df_trial[!(df_trial$group=="controlgroup"), ]
```

```{r}
# user actions for all trials, only experimental groups

df <- df[!(df$group=="controlgroup"), ]

```



## Analyse how often participants interacted with the prototype (in total and per group)

#### adding notes 

##### in total
```{r}
sum(df_trial$add_note)
sum(df_exp$add_note)
mean(df_exp$add_note)
sd(df_exp$add_note)
max(df_trial$add_note)
```
##### per group across trials
```{r}
aggregate(df_exp$add_note, by=list(group = df_exp$group), FUN=mean)
aggregate(df_exp$add_note, by=list(group = df_exp$group), FUN=sd)
```
##### per group for each trial
```{r}
aggregate(df$add_note, by=list(group = df$group, trial = df$trial), FUN=mean)
aggregate(df$add_note, by=list(group = df$group, trial = df$trial), FUN=sd)
```
##### per trial across groups
```{r}
aggregate(df$add_note, by=list(trial = df$trial), FUN=mean)
aggregate(df$add_note, by=list(trial = df$trial), FUN=sd)
```

##### Plots
```{r}
add_notes <- ggline(df, x = "trial", y ="add_note", 
       add = c("mean_se"), 
       color = "group", palette = c('darkgoldenrod2', 'darkcyan'), title = "usage of adding notes", 
       ylab = "number of uses", 
       legend = "none")
ggpar(add_notes, ylim = c(0, 9))
ggsave("plots_participants_interaction_with_prototype/add_notes.png")
```
```{r}
c <- table(df_exp$add_note)
c
barplot(height = c, 
        xlab = "nr of uses (per trial)", 
ylab = "by how many participants", 
main = "adding notes to the sidebar")
```



#### adding content to context

##### in total
```{r}
sum(df_trial$add_to_context)
mean(df_exp$add_to_context)
sd(df_exp$add_to_context)
max(df_trial$add_to_context)
```
##### per group across trials
```{r}
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=mean)
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=sd)
```
##### per group for each trial
```{r}
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=sd)
```
##### per trial across groups
```{r}
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=sd)
```
##### Plots
```{r}
add_content <- ggline(df, x = "trial", y ="add_to_context", 
       add = c("mean_se"), 
       color = "group",        
       palette = c('darkgoldenrod2', 'darkcyan')
, 
       title = "usage of adding content to the context ", 
       legend = "none", ylab = "number of uses")

ggpar(add_content, ylim = c(0, 9))

ggsave("plots_participants_interaction_with_prototype/add_content.png")
```

```{r}
c <- table(df_exp$add_to_context)
c
barplot(height = c, 
        xlab = "nr of uses (per trial)", 
ylab = "by how many participants", 
main = "adding content to the sidebar")
```


#### removing content from the context

##### in total
```{r}
sum(df_trial$remove_from_context)
mean(df_exp$remove_from_context)
max(df_trial$remove_from_context)
sd(df_exp$remove_from_context)
```


per group across trials
```{r}
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=mean)
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=sd)
```
per group for each trial
```{r}
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=sd)
```
per trial across groups
```{r}
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=sd)
```
##### Plots
```{r}
remove_content <- ggline(df, x = "trial", y ="remove_from_context", 
       add = c("mean_se"), 
       legend = "none", 
       color = "group", palette = c('darkgoldenrod2', 'darkcyan'), title = "usage of removing content ", ylab = "number of uses")

ggpar(remove_content, ylim = c(0, 9))

```

```{r}
c <- table(df_exp$remove_from_context)
c
barplot(height = c, 
        xlab = "nr of uses (per trial)", 
ylab = "by how many participants", 
main = "removing content from the context")
```


#### highlighting content

##### in total
```{r}
sum(df_trial$highlights)
mean(df_exp$highlights)
max(df_trial$highlights)
sd(df_trial$highlights)
```

##### per group across trials
```{r}
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=mean)
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=sd)
```
##### per group for each trial
```{r}
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=sd)
```
##### per trial across groups
```{r}
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=sd)
```
##### Plots
```{r}
highlight <- ggline(df, x = "trial", y ="highlights", 
       add = c("mean_se"), 
       color = "group", palette = c('darkgoldenrod2', 'darkcyan'), 
       title = "usage of highlight function "
      , ylab = "number of uses", 
       legend = "none")
ggpar(highlight, ylim = c(0, 9))

ggsave("plots_participants_interaction_with_prototype/highlight.png")
```

```{r}
c <- table(df_exp$highlights)
c
barplot(height = c, 
        xlab = "nr of uses (per trial)", 
ylab = "by how many participants", 
main = "opening websites from the sidebar")
```
##### open websites from sidebar

##### in total
```{r}
sum(df_trial$open_website_from_sidebar)
mean(df_exp$open_website_from_sidebar)
max(df_trial$open_website_from_sidebar)
sd(df_exp$open_website_from_sidebar)
```

##### per group across trials
```{r}
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=mean)
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=sd)
```
##### per group for each trial
```{r}
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=sd)
```
##### per trial across groups
```{r}
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=sd)
```
##### Plots
```{r}
open_website <- ggline(df, x = "trial", y ="open_website_from_sidebar", 
       add = c("mean_se"), 
       color = "group", 
       palette = c('darkgoldenrod2', 'darkcyan'), 
       ylab = "number of uses", title = "usage of open website from the sidebar", 
       legend = "none")

ggpar(open_website, ylim = c(0, 9))

ggsave("plots_participants_interaction_with_prototype/open_websites.png")
```
```{r}
c <- table(df_exp$open_website_from_sidebar)
c
barplot(height = c, 
        xlab = "nr of uses (per trial)", 
ylab = "by how many participants", 
main = "opening websites from the sidebar")
```


##### removing activities from the activity list

##### in total
```{r}
sum(df_trial$remove_activity)
sd(df_exp$remove_activity)
mean(df_exp$remove_activity)
max(df_trial$remove_activity)

```

##### per group across trials
```{r}
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=mean)
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=sd)
```
##### per group for each trial
```{r}
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=sd)
```

##### per trial across groups
```{r}
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=sd)
```

##### Plots
```{r}
remove_activities <- ggline(df, x = "trial", y ="remove_activity", 
       add = c("mean_se"), 
       color = "group", palette = c('darkgoldenrod2', 'darkcyan'), 
       title = "usage of removing activities from the acitivity list", 
        ylab = "number of uses", 
       legend = "none")

ggpar(remove_activities, ylim = c(0, 9))

ggsave("plots_participants_interaction_with_prototype/remove_activities.png")
```

```{r}
c <- table(df_exp$remove_activity)
c
barplot(height = c, 
        xlab = "nr of uses (per trial)", 
ylab = "by how many participants", 
main = "removing activities from the activity list")
```


##### removing notes

##### in total
```{r}
sum(df_trial$remove_note)
mean(df_exp$remove_note)
max(df_trial$remove_note)
sd(df_exp$remove_note)
```


##### per group across trials
```{r}
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=mean)
aggregate(df_exp$add_to_context, by=list(group = df_exp$group), FUN=sd)
```
##### per group for each trial
```{r}
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(group = df$group, trial = df$trial), FUN=sd)
```
##### per trial across groups
```{r}
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=mean)
aggregate(df$add_to_context, by=list(trial = df$trial), FUN=sd)
```

##### Plots
```{r}
remove_notes <- ggline(df, x = "trial", y ="remove_note", 
       add = c("mean_se"), 
       color = "group", 
       title = "usage of removing notes ", 
       palette = c('darkgoldenrod2', 'darkcyan'), 
       ylab = "number of uses", 
       legend = "none"
)

ggpar(remove_notes, ylim = c(0, 9))

ggsave("plots_participants_interaction_with_prototype/remove_notes.png")

```

```{r}
c <- table(df_exp$remove_note)
c
barplot(height = c, 
        xlab = "nr of uses (per trial)", 
ylab = "by how many participants", 
main = "removing a note from the sidebar")
```



