---
title: "System Usability Scale"
output: html_notebook
---
In this notebook, I analyze if the System Usability Scale score differs between groups and if the SUS scores are related to other variables, including NASA-TLX, performance and TA-EG. 

## Load necessary packages
```{r}
library(WRS2)
library(rcompanion)
library(feather)
library(ggplot2)  
library(plyr)
library(ez)
library(schoRsch)
library(MASS)
library(car)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(here)
```


## Load the data

We are loading the experimental data in wide format. 
```{r}
df_wide <- read_feather(path = here('../../data_frames/results_test_18-03-2020.file'))
```


## Prepare the data

### Assign participants to the experimental group 
* version 1 and 2 --> control group
* version 3 und 4 --> experimental group context free
* version 5 und 6--> experimental group context sensitive

```{r}
df_wide$group <- ifelse((df_wide$version == 5 | df_wide$version == 6), 'context-sensitive experimental group', ifelse((df_wide$version == 1 | df_wide$version == 2), 'controlgroup', 'context-free experimental group'))
```

```{r}
# Check if the design is balanced
table(df_wide$group)
```


```{r}
# Convert trial, group and subject number to factors in each dataframe

df_wide$group <- factor(df_wide$group)
df_wide$partnr <- factor(df_wide$partnr)

```

## Data Analysis: System Usability scale

#### Are there differences between the groups?
Computing Kruskal-Wallis rank sum test yields no significant differences. 
```{r}
tapply(df_wide$SUS, df_wide$group, mean)
tapply(df_wide$SUS, df_wide$group, sd)
mean(df_wide$SUS)
sd(df_wide$SUS)
kruskal.test(df_wide$SUS ~ df_wide$group)
```


#### Are there differences between the versions?
Computing Kruskal-Wallis rank sum test yields no significant differences. 
```{r}
tapply(df_wide$SUS, df_wide$version, mean)
kruskal.test(df_wide$SUS ~ df_wide$version)
```

#### How many participants in the experimental groups gave the system a "positive" rating (> 71.4)?

Only four participants rated the prorotpye above 71.4.
```{r}
df_exp <- df_wide[!(df_wide$group=="controlgroup"), ]

df_exp$SUS_above_average <- ifelse((df_exp$SUS > 71.4), 1, 0)
df_exp$SUS_above_average <- as.factor(df_exp$SUS_above_average)
df_exp %>%
  count(SUS_above_average)
```


###### boxplots
```{r}
ggplot(df_wide,aes(x = group, y=SUS) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "SUS",
       title="SUS results by group")

```

### Relation with NASA-TLX
Correlate SUS score and NASA-TLX mean (n.s.). Since the boxplots show some outliers, we compute robust correlation rather than pearson correlation. Trend: the higher the perceived system usability, the lower the perceived mental workload.
```{r}
df_wide$NASATLX_mean <-  rowMeans(df_wide[c('NASATLX_1', 'NASATLX_4', 'NASATLX_3', 'NASATLX_4')], na.rm=TRUE)
pbcor(df_wide$SUS,df_wide$NASATLX_mean)
```


There is a significant negative correlation with NASA-TLX before interruption (p = .019)
```{r}
pbcor(df_wide$SUS,df_wide$NASATLX_1)
plot(df_wide$SUS~df_wide$NASATLX_1) + abline(lm(df_wide$SUS~df_wide$NASATLX_1), col="blue")
```


Correlations with later trials are not significant.
```{r}
pbcor(df_wide$SUS,df_wide$NASATLX_2)
pbcor(df_wide$SUS,df_wide$NASATLX_3)
pbcor(df_wide$SUS,df_wide$NASATLX_4)
```


Compare the relation between NASA-TLX score before the first interruption and SUS scores between groups.
```{r}
df_control <- df_wide[(df_wide$group=="controlgroup"),]
df_exp_contextfree <- df_wide[(df_wide$group=="context-free experimental group"),]
df_exp_context <- df_wide[(df_wide$group=="context-sensitive experimental group"),]

plot(df_wide$NASATLX_1~df_wide$SUS) + abline(lm(df_control$NASATLX_1~df_control$SUS), col="blue") + abline(lm(df_exp_contextfree$NASATLX_1~df_exp_contextfree$SUS), col="green") + abline(lm(df_exp_context$NASATLX_1~df_exp_context$SUS), col="red")
```


It turns out that the correlation is mostly pronounced (and only statistically significant) for the context-sensitive experimental group. 
```{r}
pbcor(df_control$SUS,df_control$NASATLX_1)
pbcor(df_exp_contextfree$SUS,df_exp_contextfree$NASATLX_1)
pbcor(df_exp_context$SUS,df_exp_context$NASATLX_1)

# plotting NASA-TLX scores before the first interruption and SUS scores for the context-sensitive experimental group
plot(df_exp_context$NASATLX_1~df_exp_context$SUS)  + abline(lm(df_exp_context$NASATLX_1~df_exp_context$SUS), col="green")
```
When comparing the overall NASA-TLX score and the SUS scores between groups, a similar trend for the context-sensitive experimental group emerges.
```{r}
pbcor(df_control$SUS,df_control$NASATLX_mean)
pbcor(df_exp_contextfree$SUS,df_exp_contextfree$NASATLX_mean)
pbcor(df_exp_context$SUS,df_exp_context$NASATLX_mean)

plot(df_exp_context$NASATLX_mean~df_exp_context$SUS)  + abline(lm(df_exp_context$NASATLX_mean~df_exp_context$SUS), col="green")
```
#### Exploring the relation between SUS and NASA-TLX values for the context-sensitive group
For the context-sensitive experimental group, all correlations are significant: the higher the perceived usability, the lower the perceived mental workload. 
```{r}
pbcor(df_exp_context$SUS,df_exp_context$NASATLX_2)
pbcor(df_exp_context$SUS,df_exp_context$NASATLX_3)
pbcor(df_exp_context$SUS,df_exp_context$NASATLX_4)
```

#### Performance

There is no significant correlation between performance and SUS score, neither on a global nor on a group level. 
```{r}
df_wide <- transform(df_wide, perf = mistakes_after_interruption / entries_after_interruption)
pbcor(df_wide$perf, df_wide$SUS)
```

Context-sensitive experimental group
```{r}
df_exp_context <- transform(df_exp_context, perf = mistakes_after_interruption / entries_after_interruption)
pbcor(df_exp_context$perf, df_exp_context$SUS)
```


Context-free experimental group
```{r}
df_exp_contextfree <- transform(df_exp_contextfree, perf = mistakes_after_interruption / entries_after_interruption)
pbcor(df_exp_contextfree$perf, df_exp_contextfree$SUS)
```


Control group
```{r}
df_control <- transform(df_control, perf = mistakes_after_interruption / entries_after_interruption)
pbcor(df_control$perf, df_control$SUS)
```



## Resumption speed

There is no significant correlation between resumption speed  and SUS score, neither on a global nor on a group level. 
```{r}
# global level
# compute resumption speed for each trial
df_wide$res_speed_1 <- 1/df_wide$resumption_lag_1
df_wide$res_speed_2 <- 1/df_wide$resumption_lag_2
df_wide$res_speed_3 <- 1/df_wide$resumption_lag_3

# compute correlations between SUS scores and resumption speed for each trial
pbcor(df_wide$SUS,df_wide$res_speed_1)
pbcor(df_wide$SUS,df_wide$res_speed_2)
pbcor(df_wide$SUS,df_wide$res_speed_3)
```


```{r}
#Context-sensitive experimental group
# compute resumption speed for each trial

df_exp_context$res_speed_1 <- 1/df_exp_context$resumption_lag_1
df_exp_context$res_speed_2 <- 1/df_exp_context$resumption_lag_2
df_exp_context$res_speed_3 <- 1/df_exp_context$resumption_lag_3

# compute correlations between SUS scores and resumption speed for each trial

pbcor(df_exp_context$SUS,df_exp_context$res_speed_1)
pbcor(df_exp_context$SUS,df_exp_context$res_speed_2)
pbcor(df_exp_context$SUS,df_exp_context$res_speed_3)
```


```{r}
#Context-free experimental group
# compute resumption speed for each trial

df_exp_contextfree$res_speed_1 <- 1/df_exp_contextfree$resumption_lag_1
df_exp_contextfree$res_speed_2 <- 1/df_exp_contextfree$resumption_lag_2
df_exp_contextfree$res_speed_3 <- 1/df_exp_contextfree$resumption_lag_3

# compute correlations between SUS scores and resumption speed for each trial
pbcor(df_exp_contextfree$SUS,df_exp_contextfree$res_speed_1)
pbcor(df_exp_contextfree$SUS,df_exp_contextfree$res_speed_2)
pbcor(df_exp_contextfree$SUS,df_exp_contextfree$res_speed_3)
```


```{r}
#Control group
# compute resumption speed for each trial

df_control$res_speed_1 <- 1/df_control$resumption_lag_1
df_control$res_speed_2 <- 1/df_control$resumption_lag_2
df_control$res_speed_3 <- 1/df_control$resumption_lag_3

# compute correlations between SUS scores and resumption speed for each trial
pbcor(df_control$SUS,df_control$res_speed_1)
pbcor(df_control$SUS,df_control$res_speed_2)
pbcor(df_control$SUS,df_control$res_speed_3)
```
## List input lags

There is no significant correlation between list input lags and SUS score, neither on a global nor on a group level.

```{r}
# global level
# compute correlations between SUS scores and list input lags for each trial

pbcor(df_wide$SUS,df_wide$list_input_lag_1)
pbcor(df_wide$SUS,df_wide$list_input_lag_2)
pbcor(df_wide$SUS,df_wide$list_input_lag_3)
```


```{r}
# Control group
# compute correlations between SUS scores and list input lags for each trial

pbcor(df_control$SUS,df_control$list_input_lag_1)
pbcor(df_control$SUS,df_control$list_input_lag_2)
pbcor(df_control$SUS,df_control$list_input_lag_3)
```

```{r}
# context-sensitive experimental group
# compute correlations between SUS scores and list input lags for each trial

pbcor(df_exp_context$SUS,df_exp_context$list_input_lag_1)
pbcor(df_exp_context$SUS,df_exp_context$list_input_lag_2)
pbcor(df_exp_context$SUS,df_exp_context$list_input_lag_3)
```

```{r}
#context-free experimental group
# compute correlations between SUS scores and list input lags for each trial

pbcor(df_exp_contextfree$SUS,df_exp_contextfree$list_input_lag_1)
pbcor(df_exp_contextfree$SUS,df_exp_contextfree$list_input_lag_2)
pbcor(df_exp_contextfree$SUS,df_exp_contextfree$list_input_lag_3)
```


## purposeful action lags

There is no significant correlation between purposeful action lags and SUS score. 
```{r}
# df without control group
df_exp <- df_wide[!(df_wide$group=="controlgroup"),]

# convert group and trial to factors
df_exp$group <- factor(df_exp$group)

# log-transform purposeful action lags to achieve better distribution characteristics
df_exp$ppa_log_1 <- log(df_exp$purposeful_action_lag_1)
df_exp$ppa_log_2 <- log(df_exp$purposeful_action_lag_2)
df_exp$ppa_log_3 <- log(df_exp$`purposeful_action_lag.3`)


## compute correlations
cor.test(df_exp$ppa_log_1, df_exp$SUS, use='complete.obs')
cor.test(df_exp$ppa_log_2, df_exp$SUS, use='complete.obs')
cor.test(df_exp$ppa_log_3, df_exp$SUS, use='complete.obs')


```

