---
title: "TAEG"
output: html_notebook
---
In this notebook, I analyze if the technology affinity as measured by the TA-EG differs between groups and if the TA-EG scores are related to other variables, including NASA-TLX, performance and TA-EG. 


## Include necessary packages 
```{r}
library(WRS2)
library(rcompanion)
library(feather)
library(ggplot2)  
library(ez)
library(schoRsch)
library(MASS)
library(car)
library(tidyverse)
library(ggpubr)
library(rstatix)

```


## Load data

We are loading the experimental data in wide format. 

```{r}
df_wide <- read_feather(path = here('../../data_frames/results_test_18-03-2020.file'))
```




## Prepare the data

### Assign participants to the experimental group 
* version 1 and 2 --> control group
* version 3 und 4 --> experimental group context free
* version 5 und 6--> experimental group context sensitive

```{r}
df_wide$group <- ifelse((df_wide$version == 5 | df_wide$version == 6), 'context-sensitive experimental group', ifelse((df_wide$version == 1 | df_wide$version == 2), 'controlgroup', 'context-free experimental group'))
```


```{r}
# Convert trial, group and subject number to factors in each dataframe
df_wide$group <- factor(df_wide$group)
df_wide$partnr <- factor(df_wide$partnr)
```



## Data Analysis
### Are there differences between the groups?

There are no signficant differences neither for the overall TAEG value nor for the subscales. 
```{r}
df_wide %>% group_by(group) %>% summarise(taeg_groups = mean(TAEG),
            taeg_sd = (sd(TAEG)/sqrt(length(TAEG))))
```

```{r}
plot(df_wide$TAEG)

## plot TA-EG mean
ggplot(df_wide,aes(x = group, y=TAEG) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "TA-EG results",
       title="TA-EG overall value by group")

## no significant group differences
kruskal.test(TAEG ~ group, data = df_wide)
```


#### TA-EG subscale enthusiasm
```{r}
ggplot(df_wide,aes(x = group, y=taeg_enthusiasm) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "TA-EG subscale enthusiasm",
       title="TA-EG subscale enthusiasm")

kruskal.test(taeg_enthusiasm ~ group, data = df_wide)
```


#### TA-EG subscale competence
```{r}
ggplot(df_wide,aes(x = group, y=taeg_competence) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "TA-EG subscale competence",
       title="TA-EG subscale competence")

kruskal.test(taeg_competence ~ group, data = df_wide)
```


#### TA-EG subscale positive attitude
```{r}
ggplot(df_wide,aes(x = group, y=taeg_pos_attitude) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "TA-EG subscale positive attitude",
       title="TA-EG subscale postive attitude")

kruskal.test(taeg_pos_attitude ~ group, data = df_wide)
```


#### TA-EG subscale negative attitude
```{r}
ggplot(df_wide,aes(x = group, y=taeg_neg_attitude) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "TA-EG subscale negative attitude",
       title="TA-EG subscale negative attitude")

kruskal.test(taeg_neg_attitude ~ group, data = df_wide)
```


### Is there a relation between TA-EG scores and performance?


#### Correlation with nr of mistakes / entries after first interruption

```{r}
#calculate the measures

df_wide <- transform(df_wide, perf = mistakes_after_interruption / entries_after_interruption)
```


There is no significant correlation between TAEG mean and performance.
```{r}
plot(df_perf$perf~df_perf$TAEG)
cor.test(df_perf$perf, df_perf$TAEG, method = 'pearson')
```


##### TA-EG subscale enthusiasm (n.s.)



```{r}
plot(df_perf$perf~df_perf$taeg_enthusiasm)
cor.test(df_perf$perf, df_perf$taeg_enthusiasm, method = 'pearson')
```


##### TA-EG subscale competence (n.s.)


```{r}
plot(df_perf$perf~df_perf$taeg_competence)
cor.test(df_perf$perf, df_perf$taeg_enthusiasm, method = 'pearson')
```


##### TA-EG subscale positive attitude (n.s.)


```{r}
plot(df_perf$perf~df_perf$taeg_pos_attitude)
cor.test(df_perf$perf, df_perf$taeg_pos_attitude, method = 'pearson')
```


##### TA-EG subscale positive attitude (n.s.)

```{r}
plot(df_perf$perf~df_perf$taeg_neg_attitude)
cor.test(df_perf$perf, df_perf$taeg_neg_attitude, method = 'pearson')
```


## influence on NASA-TLX values


There is a significant correlation between participants' mean NASA-TLX score and their TA-EG score
```{r}
library(matrixStats)
```

#### overall NASA-TLX score
```{r}
nasa <- cbind(df_wide$NASATLX_1, df_wide$NASATLX_2, df_wide$NASATLX_3, df_wide$NASATLX_4)
df_wide$NASATLX_mean <- rowMeans2(nasa)

cor.test(df_wide$TAEG, df_wide$NASATLX_mean, method = 'pearson')
```
##### subscale enthusiasm (n.s.)
```{r}
cor.test(df_wide$taeg_enthusiasm, df_wide$NASATLX_mean, method = 'pearson')
```
subscale competence (p = .038)
```{r}
cor.test(df_wide$taeg_competence, df_wide$NASATLX_mean, method = 'pearson')
```
##### subscale positive attitude (n.s.)
```{r}
cor.test(df_wide$taeg_pos_attitude, df_wide$NASATLX_mean, method = 'pearson')
```
#### subscale negative attitude (p = .046)
```{r}
cor.test(df_wide$taeg_neg_attitude, df_wide$NASATLX_mean, method = 'pearson')
```

```{r}
par(mgp=c(2,1,0), mar=c(3,3,1,1))
# Fit regression line
require(stats)
reg<-lm(df_wide$NASATLX_mean~df_wide$TAEG)
coeff=coefficients(reg)
# equation of the line : 
eq = paste0("y = ", round(coeff[2],1), "*x ", round(coeff[1],1))
# plot
plot(df_wide$NASATLX_mean~df_wide$TAEG, main=eq, ylab = "NASA-TLX mean score", xlab = "TA-EG score", title = "Correlation between NASA-TLX and TA-EG score") + abline(reg, col="blue") 
```

#### warm up task (n.s.)

```{r}
plot(df_wide$TAEG~df_wide$NASATLX_1)
cor.test(df_wide$TAEG,df_wide$NASATLX_1, method = 'pearson')
```

##### subscale taeg_enthusiasm (n.s.)
```{r}
plot(df_wide$NASATLX_1  ~df_wide$taeg_enthusiasm)
cor.test(df_wide$taeg_enthusiasm,df_wide$NASATLX_1, method = 'pearson')
```

##### subscale taeg_competence (n.s.)
```{r}
plot(df_wide$NASATLX_1~df_wide$taeg_competence)
cor.test(df_wide$taeg_competence,df_wide$NASATLX_1, method = 'pearson')
```

##### subscale taeg_pos_attitude (n.s.) 
```{r}
plot(df_wide$NASATLX_1~df_wide$taeg_pos_attitude)
cor.test(df_wide$taeg_pos_attitude,df_wide$NASATLX_1, method = 'pearson')
```

##### subscale taeg_neg_attitude (p = .023)

```{r}
par(mgp=c(2,1,0), mar=c(3,3,1,1))
# Fit regression line
require(stats)
reg<-lm(df_wide$NASATLX_1~df_wide$taeg_neg_attitude)
coeff=coefficients(reg)
# equation of the line : 
eq = paste0("y = ", round(coeff[2],1), "*x ", round(coeff[1],1))
# plot
plot(df_wide$NASATLX_1~df_wide$taeg_neg_attitude, main=eq) + abline(reg, col="blue")


cor.test(df_wide$taeg_neg_attitude,df_wide$NASATLX_1, method = 'pearson')
```





#### first trial


no significant correlation
```{r}
plot(df_wide$TAEG~df_wide$NASATLX_2)
cor.test(df_wide$TAEG,df_wide$NASATLX_2, method = 'pearson')
```


##### subscale taeg_enthusiasm (n.s.)
```{r}
plot(df_wide$NASATLX_2  ~df_wide$taeg_enthusiasm)
cor.test(df_wide$taeg_enthusiasm,df_wide$NASATLX_2, method = 'pearson')
```


##### subscale taeg_competence (n.s.)
```{r}
plot(df_wide$NASATLX_2~df_wide$taeg_competence)
cor.test(df_wide$taeg_competence,df_wide$NASATLX_2, method = 'pearson')
```


##### subscale taeg_pos_attitude (n.s.)
```{r}
plot(df_wide$NASATLX_2~df_wide$taeg_pos_attitude)
cor.test(df_wide$taeg_pos_attitude,df_wide$NASATLX_2, method = 'pearson')
```


##### subscale taeg_neg_attitude (n.s.)
```{r}
plot(df_wide$NASATLX_2~df_wide$taeg_neg_attitude)
cor.test(df_wide$taeg_neg_attitude,df_wide$NASATLX_2, method = 'pearson')

```


#### second trial (p = .013)



significant positive correlation
```{r}
plot(df_wide$TAEG~df_wide$NASATLX_3) + abline(lm(df_wide$TAEG~df_wide$NASATLX_3), col = 'blue')
cor.test(df_wide$TAEG,df_wide$NASATLX_3, method = 'pearson')
```



##### subscale taeg_enthusiasm (n.s.)
```{r}
plot(df_wide$NASATLX_3  ~df_wide$taeg_enthusiasm)
cor.test(df_wide$taeg_enthusiasm,df_wide$NASATLX_3, method = 'pearson')
```


subscale taeg_competence (p <.01)
```{r}
plot(df_wide$NASATLX_3~df_wide$taeg_competence)
cor.test(df_wide$taeg_competence,df_wide$NASATLX_3, method = 'pearson')
```


##### subscale taeg_pos_attitude (n.s.)
```{r}
plot(df_wide$NASATLX_3~df_wide$taeg_pos_attitude)
cor.test(df_wide$taeg_pos_attitude,df_wide$NASATLX_3, method = 'pearson')
```


##### subscale taeg_neg_attitude (n.s.)
```{r}
plot(df_wide$NASATLX_3~df_wide$taeg_neg_attitude)
cor.test(df_wide$taeg_neg_attitude,df_wide$NASATLX_3, method = 'pearson')
```


#### fourth trial (p = .014)


```{r}
plot(df_wide$TAEG~df_wide$NASATLX_4) + abline(lm(df_wide$TAEG~df_wide$NASATLX_4), col = "blue")
cor.test(df_wide$TAEG,df_wide$NASATLX_4, method = 'pearson')
```
##### subscale enthusiasm (n.s.)
```{r}
plot(df_wide$NASATLX_4  ~df_wide$taeg_enthusiasm)
cor.test(df_wide$taeg_enthusiasm,df_wide$NASATLX_4, method = 'pearson')
```


##### subscale competence (p = .029)
```{r}
plot(df_wide$NASATLX_4~df_wide$taeg_competence)
cor.test(df_wide$taeg_competence,df_wide$NASATLX_4, method = 'pearson')
```


##### subscale positive attitude (n.s.)
```{r}
plot(df_wide$NASATLX_4~df_wide$taeg_pos_attitude)
cor.test(df_wide$taeg_pos_attitude,df_wide$NASATLX_4, method = 'pearson')
```


##### subscale negative attitude (p = .029)
```{r}
plot(df_wide$NASATLX_4~df_wide$taeg_neg_attitude)
cor.test(df_wide$taeg_neg_attitude,df_wide$NASATLX_4, method = 'pearson')
```


### Influence on resumption speed

```{r}
#Transform the data

df_wide$res_speed_1 <- 1/df_wide$resumption_lag_1
df_wide$res_speed_2 <- 1/df_wide$resumption_lag_2
df_wide$res_speed_3 <- 1/df_wide$resumption_lag_3
```


```{r}
# Calculate the mean resumption speed
resumption <- cbind(df_wide$res_speed_1, df_wide$res_speed_2, df_wide$res_speed_3)
df_wide$res_speed_mean <- rowMeans2(resumption)
```

```{r}
# Computing the correlation 

cor.test(df_wide$TAEG, df_wide$res_speed_mean, method = 'pearson')
```


##### first trial

```{r}
plot(df_wide$TAEG~df_wide$res_speed_1)
cor.test(df_wide$TAEG,df_wide$res_speed_1, method = 'pearson')
```




#### second trial (n.s.)
```{r}
plot(df_wide$TAEG~df_wide$res_speed_2)
cor.test(df_wide$TAEG,df_wide$res_speed_2, method = 'pearson')
```


#### third trial (n.s.)
```{r}
plot(df_wide$TAEG~df_wide$res_speed_3)
cor.test(df_wide$TAEG,df_wide$res_speed_3, method = 'pearson')
```

### influence on list input lag after the first trial (n.s.)
```{r}
cor.test(df_wide$TAEG,df_wide$list_input_lag_1)
```


### influence on list input lag after the second trial (n.s.)
```{r}
cor.test(df_wide$TAEG,df_wide$list_input_lag_2)
```

### influence on list input lag after the third trial (p = .037)
```{r}
cor.test(df_wide$TAEG,df_wide$list_input_lag_3)
plot(df_wide$TAEG ~ df_wide$list_input_lag_3)
```



## influence on purposeful action lags (n.s.)

```{r}
#prepare the data

## dataframe without control group
df_exp <- df_wide[!(df_wide$group=="controlgroup"),]

# convert group and trial to factors
df_exp$group <- factor(df_exp$group)

df_exp$ppa_log_1 <- log(df_exp$purposeful_action_lag_1)
df_exp$ppa_log_2 <- log(df_exp$purposeful_action_lag_2)
df_exp$ppa_log_3 <- log(df_exp$purposeful_action_lag.3)
```



```{r}
cor.test(df_exp$TAEG,df_exp$ppa_log_1)
cor.test(df_exp$TAEG,df_exp$ppa_log_2)
cor.test(df_exp$TAEG,df_exp$ppa_log_3)

```




