---
title: "Hypothesis analysis only for control and experimental group (to achieve a balanced data set)"
output: html_notebook
---

This notebook includes the statistical hypothesis tests excluding the context-free experimental group, since it has less test participants than the other two groups (N = 13 vs N = 21). This way, we can compare if the fact that the full dataset is unbalanced impacts the results of the hypothesis tests. 

```{r}
library(WRS2)
library(rcompanion)
library(feather)
library(ggplot2)  
library(ez)
library(schoRsch)
library(MASS)
library(car)
library(tidyverse)
library(here)
library(rstatix)
```

## Load the data

We are loading three versions of our dataframe. df_wide has one row per participant. 
```{r}
df_wide <- read_feather(path = here('../../data_frames/results_test_18-03-2020.file'))
```

df_long has one row per trial per participant (so three rows per participant for interruption 1, 2, 3). 
```{r}
df_long <- read_feather(path = here('../../data_frames/results_test_18-03-2020_messwdh.file'))
```

df_nasa has four rows per particpant, one for each nasa-tlx test
```{r}
df_nasa <- read_feather(path = here('../../data_frames/results_test_18-03-2020_messwdh_nasa.file'))
```

## Prepare the data

### Assign participants to the experimental group 
* version 1 and 2 --> control group
* version 3 und 4 --> context-free experimental group 
* version 5 und 6--> context-sensitive experimental group
```{r}
df_long$group <- ifelse((df_long$version == 5 | df_long$version == 6), 'context-sensitive experimental group', ifelse((df_long$version == 1 | df_long$version == 2), 'control group', 'context-free experimental group'))
df_nasa$group <- ifelse((df_nasa$version == 5 | df_nasa$version == 6), 'context-sensitive experimental group', ifelse((df_nasa$version == 1 | df_nasa$version == 2), 'control group', 'context-free experimental group'))
df_wide$group <- ifelse((df_wide$version == 5 | df_wide$version == 6), 'context-sensitive experimental group', ifelse((df_wide$version == 1 | df_wide$version == 2), 'control group', 'context-free experimental group'))

```

### Create balanced dataframe without the context-free experimental group


```{r}
df_long <- df_long[!(df_long$group=="context-free experimental group"),]
df_nasa <- df_nasa[!(df_nasa$group=="context-free experimental group"),]
df_wide <- df_wide[!(df_wide$group=="context-free experimental group"),]

```



## Testing Hypotheses

### Performance 
The statistical analysis of the data revealed no significant effects for either of the performance measure, confirming the analysis which included the context-free experimental group.


#### Correct entries
There is no significant effect for the number of correct entries.


```{r}
mean_correct_entries <- tapply(df_wide$correct_entries_after_interruption, df_wide$group, mean)
mean_correct_entries
```

```{r}
aggregate( correct_entries_after_interruption ~ group, df_wide, sd )
```

Test requirements for t-test

```{r}
## lineare model for correct_entries ~ groups
lm_cor_entries <- lm(df_wide$correct_entries_after_interruption~ df_wide$group)

# get studentized residuals
rstudent_cor_entries <- rstudent(lm_cor_entries)

qqnorm(rstudent_cor_entries); qqline(rstudent_cor_entries)

# Run Shapiro-Wilk test
shapiro.test(x = rstudent_cor_entries )

## test homogenity of variances
var.test(correct_entries_after_interruption~group, conf.level=0.95, data = df_wide, alternative=c("two.sided", "less", "greater"))
```

The tests indicate that the data is normally distributed with equal variances. Thus, we compute a t-test.

```{r}
ttest_correct_entries <- t.test(df_wide$correct_entries_after_interruption~ df_wide$group)
ttest_correct_entries
```


```{r}
# plot means between groups
barplot(mean_correct_entries, col = c('darkgoldenrod2', 'darkcyan'), xlab = 'group', ylab = 'Anzahl korrekter Einträge (Mittelwerte)')

# boxplots
ggplot(df_wide,aes(x = group, y= correct_entries_after_interruption) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "correct entries",
       title="correct entries by group")
```




#### Mistakes
There are no significant effects for the number of mistakes.

```{r}
mean_errors <- tapply(df_wide$mistakes_after_interruption, df_wide$group, mean)
```

##### test requirements for t-test


```{r}
lm_mistakes <- lm(df_wide$mistakes_after_interruption~ df_wide$group)

# get studentized residuals
rstudent_mistakes <- rstudent(lm_mistakes)

qqnorm(rstudent_mistakes); qqline(rstudent_mistakes)

# Run Shapiro-Wilk test
shapiro.test(x = rstudent_mistakes )

## test homogenity of variances
var.test(mistakes_after_interruption~group, conf.level=0.95, data = df_wide)
```

The QQplots indicate that the data is not normally distributed. Thus, we compute a Mann-Whitney-U-Test for two independent samples. 
```{r}
# compare between groups
utest_mistakes <- wilcox.test(df_wide$mistakes_after_interruption~ df_wide$group, paired=FALSE, conf.int=FALSE)
utest_mistakes
```


```{r}
# plot means between groups
barplot(mean_errors, col = c('darkgoldenrod2', 'darkcyan'), xlab = 'Group', ylab = 'Incorrect order list entries')

# boxplots
ggplot(df_wide,aes(x = df_wide$'group', y=df_wide$mistakes_after_interruption) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "mistakes",
       title="mistakes by group")
```


Exclude outliers
```{r}
tapply(df_wide$mistakes_after_interruption, df_wide$group, mean)
tapply(df_wide$mistakes_after_interruption, df_wide$group, sd)
exclude <- list(11, 26)
df_perf <- subset(df_wide, !(df_wide$partnr %in% exclude))

utest_mistakes_o <- wilcox.test(df_perf$mistakes_after_interruption~ df_perf$group, paired=FALSE, conf.int=FALSE)
utest_mistakes_o
```

#### Ratio of mistakes to total amount of entries
Statistical analysis reveals no significant effect.

```{r}
## nr of mistakes / correct entries
df_wide <- transform(df_wide, perf = mistakes_after_interruption / entries_after_interruption)
df_perf <- transform(df_perf, perf = mistakes_after_interruption / entries_after_interruption)
mean_perf <- tapply(df_wide$perf, df_wide$group, mean)
````

##### test requirements for t-test


```{r}
## lineare model for correct_entries ~ groups
lm_perf <- lm(df_wide$perf~ df_wide$group)

# get studentized residuals
rstudent_perf <- rstudent(lm_perf)

qqnorm(rstudent_perf); qqline(rstudent_perf)

# Run Shapiro-Wilk test
shapiro.test(x = rstudent_perf )

## test homogenity of variances
var.test(perf~group, conf.level=0.95, data = df_wide, alternative=c("two.sided", "less", "greater"))
```

The QQplots indicate that the data is not normally distributed. Thus, we compute a Mann-Whitney-U-Test for two independent samples. 

```{r}
# compare between groups
utest_perf <- wilcox.test(df_wide$perf~ df_wide$group, paired=FALSE, conf.int=FALSE)
utest_perf
```


```{r}
# plot means between groups
barplot(mean_perf, col = c('darkgoldenrod2', 'darkcyan'), xlab = 'group', ylab = 'Fehler pro richtige Einträge')

# boxplots
ggplot(df_wide,aes(x = df_wide$'group', y=df_wide$perf) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "mistakes per correct entries",
       title="mistakes per correct entries by group")

```


Analysis without participants 11 and 26 
```{r}
utest_perf <- wilcox.test(df_perf$perf~ df_perf$group)
utest_perf
```


### NASA-TLX
Does mental workload when working on the primary task differ between groups and over time?

As in the analysis of the complete dataset, the analysis reveals a significant main effect of trial but neither a main effect of group nor an interaction effect.   

##### Check requirements for ANOVA

```{r}
# convert trial, group and subject number to factors
df_nasa$trial <- factor(df_nasa$trial, levels=c(0,1,2,3), labels=c('0','1','2','3'))
df_nasa$group <- factor(df_nasa$group)
df_nasa$partnr <- factor(df_nasa$partnr)


# compute split-plots anova
sp_nasa_anova <- aov(NASATLX ~ trial*group + Error(partnr/(trial)), data=df_nasa)
summary(sp_nasa_anova)

## compute easy_anova as a check for split-plots anova
ez_nasa_anova = ezANOVA(
data = df_nasa
, dv = NASATLX
, wid = partnr
, within = trial
, between = group
, detailed = TRUE
)

anova_out(ez_nasa_anova)


# interaction
interaction.plot(df_nasa$`trial`, df_nasa$group, df_nasa$NASATLX, xlab = 'trial', ylab = 'NASA-TLX Mittelwert', trace.label = 'group', cex=1.3, pch=1, pt.cex = 1)
```






```{r}
# check normal distribution assumption by plotting residuals
aov_nasa.pr <- proj(sp_nasa_anova) 
nasa_res <- aov_nasa.pr[[3]][, "Residuals"]

qqnorm(nasa_res); qqline(nasa_res)

# Run Shapiro-Wilk test
shapiro.test(x = nasa_res )

# run Levene test for homogeneity of variance across groups
leveneTest(df_nasa$NASATLX~df_nasa$group)
```
QQplot of the residuals indicates that the assumption of normal distribution is not met --> compute robust ANOVA.


```{r}
## robust ANOVA
bwtrim(formula = NASATLX ~ group * trial, id = partnr, data = df_nasa)
```


### Resumption lags

Analysing distribution characteristics for raw, log-transformed and inverted resumption lags indicate that resumption speed (1/resumption lags) has the best distribution characteristics and shall therefore be used for the subsequent analysis. The analysis revealed significant main effects of group and trial, which confirms the results of the analysis which includes the context-free experimental group. 
```{r}
# convert trial, group and subject number to factors
df_long$trial <- factor(df_long$trial, levels=c(1,2,3), labels=c('1','2','3'))
df_long$group <- factor(df_long$group)
df_long$partnr <- factor(df_long$partnr)

#transform the data
df_long$reslags_speed <- 1/df_long$resumption_lags

## compute split-plots anova for log-transformed resumption lags
reslags_speed_anova <- aov(reslags_speed ~ trial*group + Error(partnr/(trial)), data=df_long)
summary(reslags_speed_anova)
```


```{r}

# check normal distribution assumption by plotting residuals
aov_reslags_speed.pr <- proj(reslags_speed_anova)

reslagspeed_res <- aov_reslags_speed.pr[[3]][, "Residuals"]

qqnorm(reslagspeed_res); qqline(reslagspeed_res)

# Run Shapiro-Wilk test
shapiro.test(x = reslagspeed_res )

# run levene test for homogenity of variance across groups
leveneTest(df_long$reslags_speed~df_long$group)

```




```{r}
# plot resumption speed 
# Boxplot for resumption speed per group

ggplot(df_long,aes(x = group, y=reslags_speed) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "resumption speed",
       title="resumption speed by group")

# interaction plot
interaction.plot(df_long$trial, df_long$group, df_long$reslags_speed, xlab = 'Task NR', ylab = 'resumption speed')
```

Post-hoc analyis using paired t-tests

```{r}
pwc <- df_long%>%
  pairwise_t_test(
    reslags_speed ~ trial, paired = TRUE,
    p.adjust.method = "bonferroni"
    )
pwc
```

### List input lags

Similar to the analysis of the full dataset, the analysis of list input lags between context-sensitive experimental group and control group yields no significant effects of group. Again, there is a trend that list input lags decrease over time which reaches significance in this case.
```{r}
## missing values per group
aggregate(df_long$list_input_lags, by= list(df_long$group),FUN=function(x) {sum(is.na(x))} )

## missing values per test participant
aggregate(df_long$list_input_lags, by= list(df_long$partnr),FUN=function(x) {sum(is.na(x))} )

## substitute values with median
df_list_md <- df_long  %>% select(partnr, trial, group,list_input_lags)

list_exp <- df_list_md$list_input_lags[which(df_list_md$group == "context-sensitive experimental group")]
list_control <- df_list_md$list_input_lags[which(df_list_md$group == "control group")]
md_exp <- median(list_exp, na.rm = TRUE)
md_control <- median(list_control, na.rm = TRUE)

# replace NA's with median
df_list_md$list_input_lags[is.na(df_list_md$list_input_lags) & df_list_md$group == "context-sensitive experimental group"] <- md_exp
df_list_md$list_input_lags[is.na(df_list_md$list_input_lags) & df_list_md$group == "control group"] <- md_control

## check NA's
## missing values per group
aggregate(df_list_md$list_input_lags, by= list(df_list_md$group),FUN=function(x) {sum(is.na(x))} )
```
## Are there differences in how fast test participants make the the first order list entry after an interruption between groups and trials ?

```{r}
## compute split-plots anova for relative resumption lags
inputlag_anova <- aov(list_input_lags ~ trial*group + Error(partnr/(trial)), data=df_list_md)
summary(inputlag_anova)

# check normal distribution assumption by plotting residuals
aov_inputlag.pr <- proj(inputlag_anova) 
inputlag_res <- aov_inputlag.pr[[3]][, "Residuals"]

qqnorm(inputlag_res); qqline(inputlag_res)

# Run Shapiro-Wilk test
shapiro.test(x = inputlag_res )


# run levene test for homogeneity of variance across groups
leveneTest(df_list_md$list_input_lags~df_list_md$group)
```
The tests indicate that the assumptions for ANOVA are met. 

```{r}
# plot interaction
interaction.plot(df_list_md$`trial`, df_list_md$group, df_list_md$list_input_lags, xlab = 'trial', ylab = 'list input lags', trace.label = 'group', cex=1.3, pch=1, pt.cex = 1)

# Boxplot for list input lag by group
ggplot(df_list_md,aes(x = df_list_md$'group', y=df_list_md$list_input_lags) ) +
  geom_boxplot() + 
  stat_summary(fun.y=mean,shape=1,col='red',geom='point') +
  labs(x="group",
       y = "list input lag",
       title="List input lag by group")

```

